title: "FARS Functions"
subtitle: "Akram Syed"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Introduction
The `FARSfun` package was developed as a part of course 3 -- *Building R Packages* -- of the *Mastering Software Development in R Specialization* offered by [Johns Hopkins University](https://www.jhu.edu/) through [Coursera](https://www.coursera.org/).

The `FARSfun` package facilitates in reading and plotting fatalities using Fatality Analysis Reporting System (FARS) data 
from the [National Highway Traffic Safety Administration](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

## 2. Setup
```{r setup}
library(FARSfun)
library(maps)
```

## 3. Data
There are three FARS datasets included within this package.

```{r data}
list.files(system.file("extdata", package = "FARSfun"))
```


## 4. Package Functions
### 4.1. `fars_read`
The `fars_read` function reads FARAS data and converts it to a `dplyr` data table. The file name is checked in the working directory. If it does not exist, the function is halted and a message is displayed to the console.

```{r read, eval=FALSE}
fars_read("accident_2015.csv.bz2")
```

### 4.2. `make_filename`
The `make_filename` function creates a file name `accident_[year].csv.bz2` given a valid year.

```{r makefilename}
make_filename("2013")
make_filename(2014)
```

### 4.3. `fars_read_years`
The `fars_read_years` function reads file(s) and creates `dat` data tables for the year(s) needed -- provided the years argument is valid.

```{r readyears, warning=FALSE}
fars_read_years(c("2013", 2014))
fars_read_years(2014:2015)
```

### 4.4. `fars_summarize_years`
The `fars_summarize_years` function retrieves data tables and summarizes counts by `year` and `MONTH`.

```{r summarize, warning=FALSE}
fars_summarize_years(c("2013", 2014, "2015"))
fars_summarize_years(2013:2015)
```

### 4.5. `fars_map_state`
The `fars_map_state` function plots fatal data on a map for the state and year provided.

```{r mapstate, warning=FALSE}
fars_map_state(12, 2014)
fars_map_state("1", "2015")
```

For more information, visit this package's [GitHub page](https://github.com/akram-syed/FARSfun).
